---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r load_data, results = "hide"}
if (!dir.exists("data")) {
    dir.create("data")
}
unzip("activity.zip", exdir = "data")

activity_data <- read.csv("data/activity.csv")

library(dplyr)
library(ggplot2)
library(huxtable)
```

## Histogram of total number of steps taken in a day in the raw data
```{r total_steps_per_day_histogram, fig.keep="all", fig.show="asis"}
activity_summary <- activity_data %>%
    mutate_if(is.factor, as.Date) %>%
    group_by(date) %>%
    filter(!is.na(steps)) %>%
    summarise(total_steps = sum(steps, na.rm = TRUE))

print(qplot(activity_summary$total_steps,
    geom = "histogram",
    bins = nrow(activity_summary),
    xlab = "Total Steps from Raw Data",
    ylab = "Count",
    col = I("red"),
    fill = I("green")))
```

## What is mean and median of the total number of steps taken per day in the raw data?
```{r total_steps_table, results = "asis"}
steps_table <- activity_data %>%
    mutate_if(is.factor, as.character) %>%
    group_by(date) %>%
    filter(!is.na(steps)) %>%
    summarise(mean_steps = mean(steps),
		median_steps = median(steps)) %>%
	select(Date = date, Mean = mean_steps, Median = median_steps)

as_hux(steps_table, add_colnames = TRUE) %>%
    set_bold(1,, TRUE) %>%
    set_bottom_border(1,, 1) %>%
	set_right_border(1, 1:2, 1) %>%
	set_align(,2:3, "center")
```

## What is the average daily activity pattern?
```{r average_daily_activity, results = "asis"}
activity_pattern <- activity_data %>%
    mutate_if(is.factor, as.character) %>%
    group_by(interval) %>%
    summarise(average_steps = mean(steps, na.rm = TRUE))

pattern_plot <- ggplot(data = activity_pattern, mapping = aes(interval, average_steps)) +
    geom_line() +
	theme_linedraw(base_size = 14) +
    xlab("Interval (5 min)") +
    ylab("Average Steps")
print(pattern_plot)
```

## Imputing missing values
```{r handle_missing_values}
missing_values <- count(activity_data[is.na(activity_data$steps),])

imputed_values <- activity_data %>%
    group_by(date) %>%
    summarise(average_steps = ceiling(mean(steps, na.rm = TRUE))) %>%
    mutate_if(is.numeric, funs(if_else(is.nan(.), 0.0, .))) %>%
    mutate_if(is.numeric, as.integer)

activity_tidy <- activity_data %>%
    inner_join(imputed_values, by = c("date")) %>%
    mutate(actual_steps = if_else(is.na(steps), average_steps, steps)) %>%
    select(date, interval, steps = actual_steps)
```
###There are `r missing_values` missing values.

## Histogram of total number of steps taken in a day in the tidy data
```{r tidy_total_steps_per_day_histogram, fig.keep="all", fig.show="asis"}
tidy_summary <- activity_tidy %>%
    mutate_if(is.factor, as.Date) %>%
    group_by(date) %>%
    summarise(total_steps = sum(steps))

print(qplot(tidy_summary$total_steps,
    geom = "histogram",
    bins = nrow(tidy_summary),
    xlab = "Total Steps from Tidy Data",
    ylab = "Count",
    col = I("red"),
    fill = I("blue")))
```

## What is mean and median of the total number of steps taken per day in the tidy data?
```{r tidy_steps_table, results = "asis"}
tidy_table <- activity_tidy %>%
    mutate_if(is.factor, as.character) %>%
    group_by(date) %>%
    summarise(mean_steps = mean(steps),
		median_steps = median(steps)) %>%
	select(Date = date, Mean = mean_steps, Median = median_steps)

as_hux(tidy_table, add_colnames = TRUE) %>%
    set_bold(1,, TRUE) %>%
    set_bottom_border(1,, 1) %>%
	set_right_border(1, 1:2, 1) %>%
	set_align(,2:3, "center")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r tidy_split_pattern, fig.keep="all", fig.show="asis"}
activity_split_pattern <- activity_tidy %>%
    mutate(weekday_weekend = if_else(is.weekend(date), "weekend", "weekday")) %>%
    mutate_at(vars(weekday_weekend), as.factor) %>%
    group_by(interval, weekday_weekend) %>%
    summarise(average_steps = mean(steps))

split_pattern_plot <- ggplot(activity_split_pattern) +
    geom_line(aes(x = interval, y = average_steps, color = weekday_weekend)) +
    facet_grid(weekday_weekend ~ .,
        labeller = as_labeller(c("weekday" = "Weekday", "weekend" = "Weekend"))) +
    theme_linedraw(base_size = 14) +
    theme(legend.position = "none") +
    xlab("Interval (5 min)") +
    ylab("Average Steps")

print(split_pattern_plot)
```